---
title: "Variable selection for prediction of DI"
author: Florian Tanner
date: "`r format(Sys.time())`"
output: html_document
---

# Steps for variable selection

see 10.1186/s13007-019-0508-7
Li 2019 Plant Methods

- Use all extracted traits in RF model, extract variable importance
- Fit simple model, test on opposite year - see test_important_variables.Rmd

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
.libPaths("C:/R-packages2")
rm(list=ls())
```

## Load packages

```{r,  message = FALSE}
library(tidyverse)
library(ggplot2); theme_set(theme_bw())
library(tidymodels)
library(doParallel)
```


## Set parameters for analysis

```{r}
random_seeds <- seq(1, 500, 1)
```

# Load and prepare data

## 2020

```{r}
scores_2020 <- read_csv("../data/2020/scores.csv")
hyperspec_full_2020 <- read_csv("../data/2020/hyperspec_full.csv")
lidar_2020 <- read_csv("../data/2020/lidar_traits.csv")
```

## 2022

```{r}
scores_2022 <- read_csv("../data/2022/scores.csv")
hyperspec_full_2022 <- read_csv("../data/2022/hyperspec_full.csv")
lidar_2022 <- read_csv("../data/2022/lidar_traits.csv")
```
```{r}
mod_data_2020 <- hyperspec_full_2020 |> 
  pivot_wider(id_cols = plot_id_fe, values_from = reflectance_smoothed, names_from = wavelength, names_prefix = "R") |> 
  left_join(scores_2020) |> 
  left_join(lidar_2020) |> 
  filter(treatment == "No fungicide") |> 
  select(-c(experiment, row, bay, rep, type)) 
```

```{r}
mod_data_2022 <- hyperspec_full_2022 |> 
  filter(!is.na(tray_id)) |> 
  pivot_wider(id_cols = tray_id, values_from = reflectance_smoothed, names_from = wavelength, names_prefix = "R") |> 
  left_join(scores_2022) |> 
  left_join(lidar_2022 |> 
              filter(!is.na(groundcover)) |> 
              mutate(tray_id = str_replace(tray_id, "_", "__"))) |> 
  filter(treatment == "Infected") |> 
  rename(plot_id_fe = tray_id) |> 
  select(-c(scoring_order, plots, splots, block))
```

## Fit random forest models to determine variable importance

```{r}
extract_vi_from_random_forests <- function(in_data = mod_data_2020) {
  out_list <- list() # Initialise list
  for(i in 1:length(random_seeds)){
    set.seed(random_seeds[i])
    rf_recipe <- recipe(pad ~ . , data = in_data) |> 
      update_role(plot_id_fe, new_role = "ID")|>
      update_role(treatment, new_role = "ID")|>
      update_role(id_with_controls, new_role = "ID") |> 
      step_normalize(all_predictors())  |> 
      step_center(all_predictors()) |>
      step_scale(all_predictors()) |>
      step_zv(all_predictors()) 
    
    rf_model <- rand_forest() |> 
      set_engine("ranger", importance = "permutation") |> 
      set_mode("regression")
    
    rf_workflow <- workflow() |> 
      add_recipe(rf_recipe) |> 
      add_model(rf_model)
    
    rf_fit <- rf_workflow |> 
      fit(data = in_data)
    
    ranger_obj <- extract_fit_parsnip(rf_fit)$fit
    
    vi <- data.frame(t(ranger_obj$variable.importance))
    vi$run <- i
    out_list[[i]] <- vi
    
  }
  return(out_list)
}
```

### Set up parallel processing

```{r}
all_cores <- parallel::detectCores(logical = FALSE)

library(doParallel)
cl <- makePSOCKcluster(all_cores)
registerDoParallel(cl)
clusterEvalQ(cl, .libPaths("C:/R-packages2/"))
```

```{r}
vi_list_2020 <- extract_vi_from_random_forests(in_data = mod_data_2020)
vi_list_2022 <- extract_vi_from_random_forests(in_data = mod_data_2022)
```


### Plot variable importance

```{r}
vip <- bind_rows(vi_list_2020) |> 
  mutate(year = 2020) |> 
  bind_rows(bind_rows(vi_list_2022) |> 
              mutate(year = 2022)) |> 
  pivot_longer(cols = c(-run, -year), names_to = "index", values_to = "importance") |> 
  mutate(index = toupper(index),
         sensor = case_when(index %in% c("GROUNDCOVER", "BIOMASS") ~ "Lidar - based",
                            TRUE ~ "Vegetation index"),
         index = case_when(index == "GROUNDCOVER" ~ "Ground cover",
                           index == "BIOMASS" ~ "Biomass",
                           TRUE ~ index))
```

```{r}
vip |> 
  write_csv("out/metrics/variable_importance_predict_di.csv")
```

```{r}
sessionInfo()
```
