---
title: "Trait EDA"
author: Florian Tanner
date: "`r format(Sys.time())`"
output: html_document
---


```{r}
rm(list=ls())
`%nin%` = Negate(`%in%`)
```

## Load packages

```{r,  message = FALSE}
library(tidyverse)
library(ggplot2); theme_set(theme_bw())
library(patchwork)
```

## Load data

```{r}
hyperspec_2020 <- read_csv("../data/2020/hyperspec_full.csv")
scores_2020 <- read_csv("../data/2020/scores.csv")
```
```{r}
hyperspec_2022 <- read_csv("../data/2022/hyperspec_full.csv")
scores_2022 <- read_csv("../data/2022/scores.csv")
```
```{r}
regression_mean_feature_importance <- read_csv("../predict_DI/out/metrics/mean_variable_importance_predict_di.csv")

classification_mean_feature_importance <- read_csv("../classify_treatment/out/metrics/mean_variable_importance_classify_di.csv")
```
```{r}
regression_wavelength_importance <- regression_mean_feature_importance |> 
  filter(index %nin% c("Biomass", "Ground cover")) |> 
  mutate(wavelength = as.numeric(str_remove(index, "R"))) 

regression_lidar_importance <-  regression_mean_feature_importance |> 
  filter(index %in% c("Biomass", "Ground cover")) |> 
  mutate(index = case_when(index == "Biomass" ~ "Bio- mass",
                           TRUE ~ index))

classification_wavelength_importance <- classification_mean_feature_importance |> 
  filter(index %nin% c("Biomass", "Ground cover")) |> 
  mutate(wavelength = as.numeric(str_remove(index, "R"))) 

classification_lidar_importance <-  classification_mean_feature_importance |> 
  filter(index %in% c("Biomass", "Ground cover"))|> 
  mutate(index = case_when(index == "Biomass" ~ "Bio- mass",
                           TRUE ~ index))
```

# Regression variable importance plots

## 2020

### Plot spectrum nil treatment 2020

```{r}
p_nil_spectrum_2020 <- hyperspec_2020 |> 
  left_join(scores_2020) |> 
  filter(treatment == "No fungicide") |> 
  ggplot(aes(x = wavelength, y = reflectance_smoothed, group = plot_id_fe, color = pad)) +
  geom_line(alpha = 0.7) +
  scale_color_viridis_c() +
  scale_x_continuous(breaks = seq(400, 1600, 200)) +
  theme(legend.position = "top", legend.box = "horizontal",
        legend.title = element_text(vjust = 0.8)) +
  labs(x = "Wavelength [nm]", 
       y = "Reflectance", 
       color = "Plots 2020: Percent AB damage (No Fungicide treatment only)")
```

### Variable importance regression 2020

```{r}
p_wavelength_importance_regression_2020 <- regression_wavelength_importance |> 
  filter(year == 2020) |> 
  ggplot(aes(x = wavelength, y = rescaled_importance, color = rescaled_importance, fill = rescaled_importance)) +
  geom_col() +
  scale_y_continuous(limits = c(-1,1)) +
  scale_x_continuous(breaks = seq(400, 1600, 200)) +
  scico::scale_color_scico(palette = "vik", limits = c(-1,1)) +
  scico::scale_fill_scico(palette = "vik", limits = c(-1,1)) +
  theme(legend.position = "bottom",
        legend.title = element_text(vjust = 0.8)) +
  labs(x= "Wavelength [nm]", 
       y = "VI",
       color = "Variable importance (scaled)",
       fill = "Variable importance (scaled)")
```


### Importance lidar traits regression 2020 

```{r}
p_lidar_importance_regression_2020  <-  regression_lidar_importance |> 
  filter(year == 2020) |> 
  ggplot(aes(x = index, y = rescaled_importance, fill = rescaled_importance)) +
  geom_col(width = 0.6, color = "black") +
  scico::scale_color_scico(palette = "vik", limits = c(-1,1)) +
  scico::scale_fill_scico(palette = "vik", limits = c(-1,1)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 4)) +
  scale_y_continuous(limits = c(-1,1)) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        plot.title = element_blank()) +
  labs(color = "Variable importance (scaled)",
       fill = "Variable importance (scaled)",
       y = "VI")

p_lidar_importance_regression_2020
```


### Combine regression 2020 plot

```{r}
reg_plot_2020 <- p_nil_spectrum_2020 + p_wavelength_importance_regression_2020 + 
  plot_layout(nrow = 2, 
              design = "A
                        A
                        A
                        B") + 
  inset_element(p_lidar_importance_regression_2020, 0,0,0.25,0.45, align_to = 'full')
```

## 2022

### Plot spectrum nil treatment 2022

```{r}
p_nil_spectrum_2022 <- hyperspec_2022 |> 
  left_join(scores_2022) |> 
  filter(treatment == "Infected") |> 
  ggplot(aes(x = wavelength, y = reflectance_smoothed, group = tray_id, color = pad)) +
  geom_line(alpha = 0.7) +
  scale_color_viridis_c() +
  scale_x_continuous(breaks = seq(400, 1600, 200)) +
  theme(legend.position = "top", legend.box = "horizontal",
        legend.title = element_text(vjust = 0.8)) +
  labs(x = "Wavelength [nm]", 
       y = "Reflectance", 
       color = "Trays 2022: Percent AB damage (Infected treatment only)")
```

### Variable importance regression 2022

```{r}
p_wavelength_importance_regression_2022 <- regression_wavelength_importance |> 
  filter(year == 2022) |> 
  ggplot(aes(x = wavelength, y = rescaled_importance, color = rescaled_importance, fill = rescaled_importance)) +
  geom_col() +
  scale_y_continuous(limits = c(-1,1)) +
  scale_x_continuous(breaks = seq(400, 1600, 200)) +
  scico::scale_color_scico(palette = "vik", limits = c(-1,1)) +
  scico::scale_fill_scico(palette = "vik", limits = c(-1,1)) +
  theme(legend.position = "bottom",
        legend.title = element_text(vjust = 0.8)) +
  labs(x= "Wavelength [nm]", 
       y = "VI",
       color = "Variable importance (scaled)",
       fill = "Variable importance (scaled)")
```


### Importance lidar traits regression 2022 

```{r}
p_lidar_importance_regression_2022  <-  regression_lidar_importance |> 
  filter(year == 2022) |> 
  ggplot(aes(x = index, y = rescaled_importance, fill = rescaled_importance)) +
  geom_col(width = 0.6, color = "black") +
  scico::scale_color_scico(palette = "vik", limits = c(-1,1)) +
  scico::scale_fill_scico(palette = "vik", limits = c(-1,1)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 4)) +
  scale_y_continuous(limits = c(-1,1)) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        plot.title = element_blank()) +
  labs(color = "Variable importance (scaled)",
       fill = "Variable importance (scaled)",
       y = "VI")

p_lidar_importance_regression_2022
```


### Combine regression 2022 plot

```{r}
reg_plot_2022 <- p_nil_spectrum_2022 + p_wavelength_importance_regression_2022 + 
  plot_layout(nrow = 2, 
              design = "A
                        A
                        A
                        B") + 
  inset_element(p_lidar_importance_regression_2022, 0,0,0.25,0.45, align_to = 'full')
```




# Classification variable importance plots

## 2020

### Plot spectrum nil treatment 2020

```{r}
trt_cols_2020 <- c("#5C5C5C", "#56B4E9FF")

p_all_spectrum_2020 <- hyperspec_2020 |> 
  left_join(scores_2020) |> 
  filter(treatment %in% c("No fungicide", "FN Chlorothalonil")) |> 
  ggplot(aes(x = wavelength, y = reflectance_smoothed, color = treatment)) +
  geom_line(aes(group = plot_id_fe), alpha = 0.5, size = 0.3) +
  stat_summary(fun= mean, geom = "line", size = 1, alpha =1) +
  scale_color_manual(values = trt_cols_2020) +
  scale_x_continuous(breaks = seq(400, 1600, 200)) +
  theme(legend.position = "top", legend.box = "horizontal",
        legend.title = element_text(vjust = 0.6)) +
  labs(x = "Wavelength [nm]", 
       y = "Reflectance", 
       color = "Plots 2020: Treatment")
  
p_all_spectrum_2020
```

### Variable importance classification 2020

```{r}
p_wavelength_importance_classification_2020 <- classification_wavelength_importance |> 
  filter(year == 2020) |> 
  ggplot(aes(x = wavelength, y = rescaled_importance, color = rescaled_importance, fill = rescaled_importance)) +
  geom_col() +
  scale_y_continuous(limits = c(-1,1)) +
  scale_x_continuous(breaks = seq(400, 1600, 200)) +
  scico::scale_color_scico(palette = "vik", limits = c(-1,1)) +
  scico::scale_fill_scico(palette = "vik", limits = c(-1,1)) +
  theme(legend.position = "bottom",
        legend.title = element_text(vjust = 0.8)) +
  labs(x= "Wavelength [nm]", 
       y = "VI",
       color = "Variable importance (scaled)",
       fill = "Variable importance (scaled)")
```


### Importance lidar traits classification 2020 

```{r}
p_lidar_importance_classification_2020  <-  classification_lidar_importance |> 
  filter(year == 2020) |> 
  ggplot(aes(x = index, y = rescaled_importance, fill = rescaled_importance)) +
  geom_col(width = 0.6, color = "black") +
  scico::scale_color_scico(palette = "vik", limits = c(-1,1)) +
  scico::scale_fill_scico(palette = "vik", limits = c(-1,1)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 4)) +
  scale_y_continuous(limits = c(-1,1)) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        plot.title = element_blank()) +
  labs(color = "Variable importance (scaled)",
       fill = "Variable importance (scaled)",
       y = "VI")

p_lidar_importance_classification_2020
```


### Combine classification 2020 plot

```{r}
class_plot_2020 <- p_all_spectrum_2020 + p_wavelength_importance_classification_2020 + 
  plot_layout(nrow = 2, 
              design = "A
                        A
                        A
                        B") + 
  inset_element(p_lidar_importance_classification_2020, 0,0,0.25,0.45, align_to = 'full')
```

## 2022

### Plot spectrum nil treatment 2022

```{r}
trt_cols_2022 <- c("#CC79A7FF", "#E69F00FF","#009E73FF")

p_all_spectrum_2022 <- hyperspec_2022 |> 
  left_join(scores_2022) |> 
  filter(!is.na(treatment)) |> 
  ggplot(aes(x = wavelength, y = reflectance_smoothed, color = treatment)) +
  geom_line(aes(group = tray_id), alpha = 0.7, size = 0.3) +
  stat_summary(fun= mean, geom = "line", size = 1, alpha =1) +
  scale_color_manual(values = trt_cols_2022) +
  scale_x_continuous(breaks = seq(400, 1600, 200)) +
  theme(legend.position = "top", legend.box = "horizontal",
        legend.title = element_text(vjust = 0.6)) +
  labs(x = "Wavelength [nm]", 
       y = "Reflectance", 
       color = "Trays 2022: Treatment")
  
p_all_spectrum_2022
```

### Variable importance classification 2022

```{r}
p_wavelength_importance_classification_2022 <- classification_wavelength_importance |> 
  filter(year == 2022) |> 
  ggplot(aes(x = wavelength, y = rescaled_importance, color = rescaled_importance, fill = rescaled_importance)) +
  geom_col() +
  scale_y_continuous(limits = c(-1,1)) +
  scale_x_continuous(breaks = seq(400, 1600, 200)) +
  scico::scale_color_scico(palette = "vik", limits = c(-1,1)) +
  scico::scale_fill_scico(palette = "vik", limits = c(-1,1)) +
  theme(legend.position = "bottom",
        legend.title = element_text(vjust = 0.8)) +
  labs(x= "Wavelength [nm]", 
       y = "VI",
       color = "Variable importance (scaled)",
       fill = "Variable importance (scaled)")
```


### Importance lidar traits classification 2022 

```{r}
p_lidar_importance_classification_2022  <-  classification_lidar_importance |> 
  filter(year == 2022) |> 
  ggplot(aes(x = index, y = rescaled_importance, fill = rescaled_importance)) +
  geom_col(width = 0.6, color = "black") +
  scico::scale_color_scico(palette = "vik", limits = c(-1,1)) +
  scico::scale_fill_scico(palette = "vik", limits = c(-1,1)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 4)) +
  scale_y_continuous(limits = c(-1,1)) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        plot.title = element_blank()) +
  labs(color = "Variable importance (scaled)",
       fill = "Variable importance (scaled)",
       y = "VI")

p_lidar_importance_classification_2022
```


### Combine classification 2022 plot

```{r}
class_plot_2022 <- p_all_spectrum_2022 + p_wavelength_importance_classification_2022 + 
  plot_layout(nrow = 2, 
              design = "A
                        A
                        A
                        B") + 
  inset_element(p_lidar_importance_classification_2022, 0,0,0.25,0.45, align_to = 'full')
```

# Save plots

```{r}
ggsave(plot= reg_plot_2020, filename = "out/graphs/reg_2020_spectrum_importance.pdf", device = "pdf", units = "cm", width = 16, height = 16, limitsize = F)

ggsave(plot= reg_plot_2022, filename = "out/graphs/reg_2022_spectrum_importance.pdf", device = "pdf", units = "cm", width = 16, height = 16, limitsize = F)

ggsave(plot= class_plot_2020, filename = "out/graphs/class_2020_spectrum_importance.pdf", device = "pdf", units = "cm", width = 16, height = 16, limitsize = F)

ggsave(plot= class_plot_2022, filename = "out/graphs/class_2022_spectrum_importance.pdf", device = "pdf", units = "cm", width = 16, height = 16, limitsize = F)
```



```{r}
sessionInfo()
```
